---
title: "Variable transformations"
author: "Michiel van Eldik"
date: "`r Sys.Date()`"
output:
   rmdformats::downcute:
    self_contained: true
    thumbnails: true
    lightbox: true
    gallery: false
    highlight: tango
---

__<a href="index.html">Back to main</a>__

# __1. Introduction__
Nothing spectacular is going on in this section.
All we're doing is transforming and renaming some variables to prepare for analyses.

```{r message=FALSE, warning=FALSE}
library(lubridate) # For dates
library(dplyr)
library(CORElearn) # For discretization
library(arulesCBA) # For discretization

```

# __2. Time-related variables__ 
This chunk changes the columns with date to an actual data format, which can be used to apply arithmetic operations. 
```{r eval = FALSE}
# To date format
brazil_df <- brazil_df %>%
  mutate(
    review_creation_date = as.Date(review_creation_date,
                                      format = "%Y-%m-%d %H:%M:%S"),
    review_answer_timestamp = as.Date(review_answer_timestamp,
                                      format = "%Y-%m-%d %H:%M:%S"),
    order_purchase_timestamp = as.Date(order_purchase_timestamp, 
                                      format = "%Y-%m-%d %H:%M:%S"),
    order_approved_at = as.Date(order_approved_at,
                                      format = "%Y-%m-%d %H:%M:%S"),
    order_delivered_carrier_date = as.Date(order_delivered_carrier_date,
                                      format = "%Y-%m-%d %H:%M:%S"),
    order_delivered_customer_date = as.Date(order_delivered_customer_date,
                                      format = "%Y-%m-%d %H:%M:%S"),
    order_estimated_delivery_date = as.Date(order_estimated_delivery_date,
                                      format = "%Y-%m-%d %H:%M:%S")
        )
```

Calculate time length through temporal differences.
```{r eval = FALSE}
brazil_df <- brazil_df %>%
  mutate(
    diff_est_deliv = order_estimated_delivery_date - order_delivered_customer_date,
    diff_pur_est  = order_estimated_delivery_date - order_purchase_timestamp,
    diff_pur_deliv = order_delivered_customer_date - order_purchase_timestamp,
    diff_rev_crea_ans = review_creation_date - review_answer_timestamp,
    diff_rev_est_ans = order_estimated_delivery_date - review_answer_timestamp,
    diff_rev_deliv_ans = order_delivered_customer_date - review_answer_timestamp
        )
```
Record temporal data.
```{r eval = FALSE}
brazil_df <- brazil_df %>%
  mutate(
    # show weekday of a date
    review_sent_dow = wday(review_creation_date, label = TRUE),
    review_answer_dow = wday(review_answer_timestamp, label = TRUE),
    # Dummy variable weekend yes/no
    review_sent_wknd = ifelse(
      review_sent_dow == "zo" | review_sent_dow == "za", 1, 0),
    review_answer_wknd = ifelse(
      review_answer_dow == "zo" | review_answer_dow == "za", 1, 0),
    # Month for seasonality
    review_sent_moy = month(review_creation_date, label = TRUE)
        )
```

```{r eval = FALSE}
# Year as seperate variables
brazil_df<- brazil_df %>%
  mutate(
    y_2016 = ifelse(year(order_purchase_timestamp) == "2016", 1, 0),
    y_2017 = ifelse(year(order_purchase_timestamp) == "2017", 1, 0),
    y_2018 = ifelse(year(order_purchase_timestamp) == "2016", 1, 0)
        )
# Year as 1 factor variable 
brazil_df$year <- "2016"
brazil_df <- brazil_df %>%
  mutate(year = ifelse(y_2017 == 1, "2017", year),
         year = ifelse(y_2018 == 1, "2018", year),
         year = as.factor(year))

brazil_df <- brazil_df %>%
  mutate(sent_sun = ifelse(review_sent_dow == "zo", 1, 0),
         sent_mon = ifelse(review_sent_dow == "ma", 1, 0),
         sent_tue = ifelse(review_sent_dow == "di", 1, 0),
         sent_wed = ifelse(review_sent_dow == "wo", 1, 0),
         sent_thu = ifelse(review_sent_dow == "do", 1, 0),
         sent_fri = ifelse(review_sent_dow == "vr", 1, 0),
         sent_sat = ifelse(review_sent_dow == "za", 1, 0)
        )
```

# __3. Message and title derived variables__

```{r eval = FALSE}
# Create dummy for message vs no message [BASED ON AFTER LEMMA]
brazil_df <- brazil_df %>%
  mutate(review_comment_message = as.character(review_comment_message),
         message_length = nchar(review_comment_message),
         message_bool = ifelse(is.na(review_comment_message), 0, 1)
        )

# Create dummy for message vs no message [BASED ON BEFORE LEMMA]
brazil_df <- brazil_df %>%
  mutate(bef_message_bool = ifelse(bef_nchar == 0, 0, 1))

# Create a dummy variable for title message
brazil_df <- brazil_df %>%
  mutate(review_comment_title = as.character(review_comment_title),
         title_length = nchar(review_comment_title),
         title_bool = ifelse(is.na(review_comment_title), 0, 1)
        )

# Create a dummy variable for title OR, AND message
brazil_df <- brazil_df %>%
  mutate(title_or_message = ifelse(message_bool == 1 | title_bool == 1, 1, 0),
         title_and_message = ifelse(message_bool == 1 & title_bool == 1, 1, 0),
         title_nor_message = ifelse(message_bool == 0 & title_bool == 0, 1, 0)
        )

# Top2box transformation
brazil_df <- brazil_df %>%
  mutate(top2box = ifelse(review_score > 3, 1, 0))
```

# __4. Region variables__
```{r eval = FALSE}
c_north <- c("AC","AP","AM","PA", "RO", "RR", "TO")
c_south <- c("SC", "RS", "PR")
c_southeast <- c("SP", "RJ", "MG", "ES")
c_northeast <- c("AL", "BA", "CE", "MA", "RN", "SE", "PI", "PB", "PE")
c_centerwest <- c("MT", "MS", "GO", "DF")

brazil_df$north <- 0
brazil_df$south <- 0
brazil_df$southeast <- 0
brazil_df$northeast <- 0
brazil_df$centerwest <- 0

brazil_df <- brazil_df %>% 
  mutate(north = ifelse(customer_state %in% c_north, 1,0),
         south = ifelse(customer_state %in% c_south, 1,0),
         southeast = ifelse(customer_state %in% c_southeast, 1,0),
         northeast = ifelse(customer_state %in% c_northeast, 1,0),
         centerwest = ifelse(customer_state %in% c_centerwest, 1,0),
        )

brazil_df <- brazil_df %>% 
  mutate(region = ifelse(north == 1, "north", ""),
         region = ifelse(northeast == 1, "northeast", region),
         region = ifelse(centerwest == 1, "centerwest", region),
         region = ifelse(southeast == 1, "southeast", region),
         region = ifelse(south == 1, "south", region),
         region = ifelse(customer_state == "DF", "southeast", region), # Belongs to southeast, culturally
         region = as.factor(region))
```

# __5. Distinguish freight-related messages__
```{r eval = FALSE}
# Due to lemmatization we don't need to worry about tenses
listje <- c("receb", # received
            "aguar", # wait 
            "ainda", # yet 
            "faltou", # missed
            "faltar", # missed 
            "incompleto", # incomplete 
            "nunca chegar", # never came / arrived
            "chegar", # To arrive
            "entregar", # deliver 
            "nao entregar", # not delivered 
            "antar do prazo", # before the term / deadline
            "prazo" # term / deadline
           )

# To fill up
brazil_df$freight_issue_bool <- 0

# If any of the words in the list can be found, add "1"
for (i in listje){
  brazil_df$freight_issue_bool <- ifelse(
    grepl(i, brazil_df$message_and_title), 
    1, 
    brazil_df$freight_issue_bool)
}

# Sanity test to see if it worked
test_out <- brazil_df[
  brazil_df$freight_issue_bool == 1, 
  c("review_comment_message",
    "freight_issue_bool",
    "review_score"
    )]

# When order is delivered after the estimated date, we name it "order_issue"
brazil_df <- brazil_df %>%
  mutate(other_issue = ifelse(diff_est_deliv > 1, 1, 0))
```

__<a href="index.html">Back to main</a>__

